---
title: "Title TBD"
bibliography: CogSci2022-register.bib
csl: apa6.csl
document-params: "10pt, letterpaper"

author-information: > 
    \author{{\large \bf Kennedy Casey} \\ University of Chicago \\ \texttt{kbcasey@uchicago.edu}
    \And {\large \bf Marisa Casillas} \\ University of Chicago \\ \texttt{mcasillas@uchicago.edu}}

abstract: >
  Child-directed speech (CDS) features words such as _doggy_, _night-night_, and _tummy_ that are rarely used in adult-directed speech (ADS). Characterisitcs of CDS word forms, such as reduplication and diminutivization, explain why they may be learned and produced earlier by children. However, it is not yet clear how or when children switch to using ADS equivalents—_dog_, _goodnight_, _stomach_. Through analysis of transcripts from CHILDES and the Language Development Project corpus, we show that children significantly increase their production of ADS word forms across age, with the average CDS-to-ADS transition point at 2.5 years. Many of the linguistic features that distinguish CDS vs. ADS registers (e.g., speech rate, lexical complexity, etc.) similarly differentiated the local speech contexts surrounding CDS vs. ADS word forms. Learners may therefore be able to capitalize on these cues to support their discovery of context-appropriate CDS/ADS pair use.
  
keywords: >
    child-directed speech; word production; linguistic input; social register; corpus analysis; developmental change
    
output: cogsci2016::cogsci_paper
final-submission: \cogscifinalcopy
---

```{r global_options, include=FALSE}
knitr::opts_chunk$set(fig.width = 3, fig.height = 3, fig.crop = F, 
                      fig.pos = "tb", fig.path = 'figs/',
                      echo = F, warning = F, cache = F, 
                      message = F, sanitize = T)
```

```{r, libraries}
library(png)
library(grid)
library(xtable)
library(kableExtra)
library(tidyverse)
library(ggpubr)
library(DescTools)

# read in outputs of "get-freq.R"
metadata <- read_csv("../../data-prep/overall/metadata.csv")
freq <- read_csv("../../data-prep/overall/freq.csv")

cds_forms <- read_csv("../../data-prep/overall/item-info.csv") %>%
    filter(form == "CDS") %>%
    pull(word)

ads_forms <- read_csv("../../data-prep/overall/item-info.csv") %>%
    filter(form == "ADS") %>%
    pull(word)

# preset colors for linguistic predictors
input_colors <- c("Prosodic" = "#1C9E78", 
                  "Lexical" = "#D95F06", 
                  "Syntactic" = "#7570B4")
```

# Introduction
Across many cultures and languages, speech that is addressed to children sounds remarkably different from speech that is addressed to adults (REFS). When communicating with young children, adults often modify their speech in ways that draw children's attention and support their several aspects of their language learning [e.g., @nencheva2021moment; @rowe2008child; @shneidman2012language; @weisleder2013talking]. As a result, child-directed speech (CDS) is differentiated from adult-directed speech (ADS) at multiple linguistic levels, including prosodic, lexical, and syntactic [@soderstrom2007beyond].

## Linguistic features of child-directed speech
CDS is associated with higher overall pitch as well as greater variability in pitch contours [@fernald1989intonation; @vosoughi2012longitudinal]. Vowels (REF) and words (REF) in CDS are often lengthened, creating utterances that are produced more slowly (REF) and that also include more pauses (REF). At the lexical level, CDS typically includes more repetition [@henning2005maternal], more words that children already know [@foushee2016lexical], and many register-specific words (e.g., in English, _doggy_, *night-night*, and _tummy_, REFS). Syntactically, CDS is characterized as less complex than ADS. CDS utterances are typically shorter (REF) and feature simpler constructions (REF). <!--kc: maybe something here about CDS being typed as a more simplified register but that's not necessarily the case. more phonetic variability, maybe another example, and seemingly redundant lexical items. the focus of the current investigation is this last point-->

[connecting sentence here]. Even very young infants appear to pick up on differences between CDS and ADS, as evidenced by their reliable preference for listening to CDS over ADS [@cooper1990preference; @manybabies2020quantifying]. In addition to this overall preference for CDS, accumulating evidence points to children's ability to distinguish individual linguistic features (REFS).

## Social context of child-directed speech
<!--kc: maybe a note here about other distinctions (e.g., acoustic and interactional). ep's paper on timbre and more about the social, oftentimes dyadic, context--> Linguistic cues are not the only features that differentiate CDS and ADS. For instance...examples like acoustic, maybe affective (if there's a ref that separates this from prosodic), and, most notably, social [@golinkoff2015baby]. BLABLABLA kids are linking linguistic and social information. Infants as young as 12 to 20 months show an emerging understanding of the relationship between registers and addressees [@soley2020infants; @ikeda2018sensitivity]. Put simply, CDS should be directed to young children, not adults [possibly a few other examples here]

## Current investigation
Here, we focus on the question of whether children may also be able to learn more subtle associations between words and registers. <!--kc: need to be careful here because we don't have data to fully speak to kids' understanding--> We return to examination of a specific lexical modification of CDS: the inclusion of register-specific words. [add stuff about why 2 labels initally seem counterintuitive but then characteristics of CDS forms that explain away the learnability issue]. Existing research does well to provide evidence for why CDS forms make up a large proportion of directed input and why they may be learned and produced earlier by children. However, conventions of adult speech <!--kc: definitely need a better phrase here--> require children to eventually transition to using ADS forms---a learning problem that has not yet received empirical attention. [connecting sentence here] <!--kc: something about the focus on lots of lang dev research on increasing vocabulary size but not necessarily context-appropriate usage or the 'loss' of words in a way-->

In Study 1, we analyze existing speech corpora for children up to age 7 to determine when children shift from CDS to ADS vocabulary [explain what we mean by this]. 

After establishing that a shift towards production of ADS forms (over CDS forms) occurs in early childhood, we explore what information in children's linguistic input could support this shift in Study 2. We investigate whether many of the linguistic features that distinguish _between_ CDS and ADS at the register level also differentiate CDS and ADS words at the utterance level. <!--kc: wording needs work here but the point that we're mostly looking at CDS is crucial-->

<!--
overview of previous work, which has largely focused on the learnability question or why these words appear in the first place
seemingly a challenge for word learners that now have to make 2 form mappings to the same meaning but diminutive suffix as a decently reliable cue to word ending [@kempe2005diminutives] 
diminutivization and reduplication predictive of learning [@ota2018choo]
existing research does well to account for why CDS forms make up a large proportion of the input and why they are learned and produced earlier by children
however, 'conventional'  (definitely need a better word here) word usage requires them to eventually transition to using ADS forms- a problem that has not yet been addressed

from there, we sought to determine which regularities might be sufficient to allow children to learn the appropriate distinctions between usage contexts for CDS vs. ADS forms

Ideas that don't really have a place yet:
mutual exclusivity--seems more tangential right now but could serve as motivation for wanting to study these words, in particular
dev changes in other input measures (e.g., overall rates of CDS, other word usage, etc.)--could motivate the longitudinal look we take here
fine-tuned input--probably not relevant unless we look at adult shifts from CDS to ADS too, later on can look at changes in the input near the transition point

CDS feature refs:
[@spinelli2017prosody]

Multiple label refs:
[@clark1987principle]

Social refs:
[@kinzler2009accent]
[@weatherhead2021toddlers]
[@weatherhead2021putting]: ME is about social inference
[@ramirez-esparza2014look]: CDS helps word learning but it's also about the social context (dyadic interaction)

Dev change/fine tuning refs:
[@ramirez-esparza2017look]: similar ideas to above but longitudinal
[@bergelson2019north]: increase in prop CDS over the first 20 months, driven by a decrease in ODS
[@jiangURexploring]: distributional semantics
[@odijk2020fine]: MLU decreasing before child word production
[@tippenhauer2020scope]: less rep reduction of low-frequency words in CDS
-->

# Study 1: When do children shift from CDS to ADS vocabulary?
Mini intro here

# Method

## Corpora

We analyzed `r metadata$n_transcripts` transcripts in the North American English collection of the Child Language Data Exchange System (CHILDES) database [@macwhinney2000childes] for children up to 7 years of age. The included transcripts were drawn from `r metadata$n_corpora` individual corpora and featured `r metadata$n_kids` children (age range = `r metadata$min_age`--`r metadata$max_age` months, _M_ = `r metadata$mean_age` months). Child production data from the Language Development Project (LDP) longitudinal corpus were also analyzed [see @rowe2008child; @huttenlocher2010sources for further details regarding participating families, recording procedures, and transcription]. These included 622 transcripts from 59 English-learning children recorded every 4 months for approximately 90 minutes from age 14 to 58 months.

## Target words
Fifteen CDS/ADS word pairs (30 total target words; see Table \ref{tab:tab1}) were selected based on (a) the appearance of at least one form on the MacArthur-Bates Communicative Development Inventory [@fenson1994variability] and (b) frequency of occurrence in CHILDES (at least 100 child-produced tokens and 100 other-produced tokens per form). Pairs were chosen such that the same object, animal, or routine could be reasonably labeled with either form (CDS or ADS) in typical communicative interactions with young children (e.g., _doggy_ or _dog_).

```{r freq_table, results = "asis"}
cds_freq <- freq %>%
    filter(word %in% cds_forms)

ads_freq <- freq %>%
    filter(word %in% ads_forms)

freq <- bind_cols(cds_freq, ads_freq) %>%
    mutate(pair = paste0(word...1, '/', word...4)) %>%
    select(pair, child...2, other...3, child...5, other...6) 
colnames(freq) <- c("Pair", "Child", "Other", "Child", "Other")

xtable::xtable(freq, digits = 0,
                             include.rownames = FALSE, 
                             label = "tab:tab1",
                             caption = "CHILDES frequency for 15 CDS/ADS word pairs. Child-produced
                             counts include tokens produced only by the target child. All other 
                             speakers' productions are included in the other-produced counts.") %>%
    xtable2kable(booktabs = TRUE, escape = FALSE) %>%
    kable_styling(latex_options = "scale_down") %>%
    add_header_above(c(" " = 2, "CDS tokens by speaker" = 2, "ADS tokens by speaker" = 2), bold = TRUE) %>%
    row_spec(0, bold = TRUE) %>%
    column_spec(2, italic = TRUE)
```

# Results
```{r}
# read in lmer outputs generated in shift-timing.R 
overall.shift <- read_csv("../../analysis/model-outputs/shift-timing.csv")

# get count of items with sig increase in ADS with age
bypair.shift <- read_csv("../../analysis/model-outputs/bypair-shift-timing.csv") %>%
  filter(p.value < 0.05) %>%
  nrow()
```
We first asked when CDS forms are replaced by ADS forms in children's own speech. We fit a mixed-effects binomial logistic regression model predicting children's production of CDS vs. ADS forms, with target child age (in months, scaled) as a single fixed effect. Random slopes and intercepts for word pairs were also included. For each target word token, the form was coded as either 0 (CDS) or 1 (ADS). Thus, the model captures, for each age, the relative proportion of CDS vs. ADS forms in children's own speech.

Children's production of ADS forms increased with age ($\beta$ = `r format(round(overall.shift$estimate, 2), 2)`, _SE_ = `r format(round(overall.shift$std.error, 2), 2)`, _t_ = `r format(round(overall.shift$statistic, 2), 2)`, _p_ `r ifelse(overall.shift$p.value < .001, paste("< 0.001"), paste("=", format(round(overall.shift$p.value, 3), 3)))`). This trend of increasing ADS form production was significant for `r bypair.shift` of 15 pairs. The average CDS-to-ADS transition point (i.e., the point at which ADS forms were produced >50% of the time) was between 24 and 36 months of age (Figure 1).

```{r shift-timing-fig, fig.env = "figure", fig.pos = "h", fig.align = "center", fig.width = 2.9, fig.height = 3.4, fig.cap = "Model-predicted increase in production of ADS forms with age. Gray lines depict individual word-pair trajectories."}

shift.timing <- png::readPNG("../figs/shift-timing.png")
grid::grid.raster(shift.timing)
```

```{r shift-timing-bypair-fig, fig.env = "figure", fig.pos = "!ht", fig.align = "center", fig.width = 3.4, fig.height = 5, fig.cap = "Inidividual word-pair trajectories for increasing production of ADS forms (blue) and descreasing production of CDS forms (red) with age. Columns indicate categorically different shift trajectory types."}

bypair.shift.timing <- png::readPNG("../figs/bypair-shift-timing.png")
grid::grid.raster(bypair.shift.timing)
```

# Study 2: What linguisitc information in children's input supports the shift from CDS to ADS vocabulary?
```{r model-plot, fig.env = "figure*", fig.pos = "h", fig.align = "center", fig.width = 7,  set.cap.width = TRUE, num.cols.cap = 2, fig.cap = "Coefficient estimates for linguistic predictors of form. Positive main effects indicate that utterances are more likely to contain ADS forms when they have higher values for the predictor (e.g., faster speech rates). Positive age interactions indicate an increasing effect of the predictor with age. Error bars depict standard errors of the coefficient estimates, and filled circles represent significant effects (\\textit{p} $<$ 0.05)."}

path <- "../../analysis/model-outputs/input/"
filenames <- list.files(path, "*.csv")
files <- lapply(paste0(path, filenames), read_csv)
input.models <- bind_rows(files, .id = "source") %>%
  rename(Predictor = term, 
         Estimate = estimate, 
         SE = std.error) %>%
  group_by(source) %>%
  mutate(order = row_number(), 
         type = case_when(
           order == 1 ~ "intercept", 
           order == 2 ~ "main_effect", 
           order == 3 ~ "age", 
           order == 4 ~ "interaction"), 
         Sig = ifelse(p.value < .05, "sig", "not_sig"), 
         Predictor = str_to_sentence(str_replace(str_remove_all(Predictor, "_scaled"), "_", " "), "first"), 
         level = ifelse(str_detect(Predictor, "Length|Verbs"), "Syntactic", 
                        ifelse(str_detect(Predictor, "Complexity|Rarity"), "Lexical", "Prosodic"))) %>%
  mutate(Predictor = factor(Predictor, levels = c("Verbs", "Verbs:age", "Length", "Length:age", "Rarity", "Rarity:age", "Complexity", "Complexity:age", "Rate", "Rate:age", "Pitch range", "Pitch range:age", "Pitch mean", "Pitch mean:age")))

main.effects <- ggplot(filter(input.models, type == "main_effect"), 
       aes(x = Estimate, y = Predictor, color = level)) +
  geom_vline(xintercept = 0, size = 0.75, linetype = "dotted", 
             color = "gray", alpha = 0.5) +
  geom_errorbar(aes(xmin = Estimate - SE, xmax = Estimate + SE), 
                width = 0, size = 0.75) + 
  geom_point(aes(shape = Sig), size = 1.5) +
  scale_shape_manual(values = c(21, 19)) +
  scale_color_manual(values = input_colors) +
  scale_x_continuous(limits = c(-0.5, 0.5)) +
  labs(title = "Main Effect", x = "Coefficient Estimate") +
  theme_test(base_size = 10) +
  theme(legend.position = "none", plot.title = element_text(hjust = 0.5))
  
interactions <- ggplot(filter(input.models, type == "interaction"), 
       aes(x = Estimate, y = Predictor, color = level)) +
  geom_vline(xintercept = 0, size = 0.5, linetype = "dotted", 
             color = "gray", alpha = 0.75) +
  geom_errorbar(aes(xmin = Estimate - SE, xmax = Estimate + SE), 
                width = 0, size = 0.75) + 
  geom_point(aes(shape = Sig), size = 1.5) +
  scale_shape_manual(values = c(21, 19), guide = "none") +
  scale_color_manual(values = input_colors) +
  scale_x_continuous(limits = c(-0.5, 0.5)) +
  labs(title = "Interaction with Age", x = "Coefficient Estimate") +
  theme_test(base_size = 10) +
  theme(legend.position = "right", legend.title = element_blank(), 
        plot.title = element_text(hjust = 0.5), 
        axis.text.y = element_blank(), axis.title.y = element_blank(), 
        axis.ticks.y = element_blank())

model.fig <- ggarrange(main.effects, interactions, widths = c(3, 3.2))
ggsave("../figs/model-outputs.png", model.fig, height = 3, width = 5, dpi = 300)

img <- png::readPNG("../figs/model-outputs.png")
grid::grid.raster(img)
```

We next explored children's input (i.e., other-produced speech), asking whether the linguistic features that differentiate CDS vs. ADS at the register level also differentiate the local speech contexts surrounding CDS vs. ADS forms. In other words, can form be predicted on the basis of individual utterance-level prosodic, lexical, or syntactic cues?

# Method

## Corpora
Reference above but drop LDP

## Linguistic input predictors

All input analyses were conducted over individual other-produced utterances in CHILDES (i.e., speech not produced by the target child). <!--kc: might add percentage of utterances from MOT/FAT just so it's clear that this is primarily adult-produced speech, not speech from other children.--> We quantified prosodic, lexical, and syntactic information to describe each utterance containing one of the 30 target words.

### Prosodic level
We measured three types of prosodic information: __mean pitch__ (Hz), __pitch range__ (Hz), and __speech rate__ (words per second). These measures were calculated over all timestamped utterances in CHILDES (`r metadata$other_timestamped`% of other-produced utterances). Utterances shorter than 58 ms were excluded from analysis. This lower bound was set by identifying the the shortest possible duration of an utterance containing at least one word in four manually annotated North American English corpora in HomeBank [@homebank; @bergelsoncorpus; @soderstromcorpus; @vandamcorpus; @warlaumontcorpus] <!--kc2mc: are these the right refs?-->. Pitch information was extracted using Praat software [@boersma2016praat].

### Lexical level
We measured two types of lexical information: complexity and rarity. __Lexical complexity__ was defined as the negative log proportion of known words in each utterance [consistent with @foushee2016lexical; @kidd2012goldilocks]. A word was considered 'known' if the age of acquisition (AoA) estimate [@fenson1994variability; @frank2017wordbank] was less than or equal to the age of the target child when they heard the utterance. Utterances with proportionally fewer known words are more lexically complex. __Lexical rarity__ was determined based on overall frequency in CHILDES. For all words with at least 10 tokens^[Manual checks revealed that many of the lowest-frequency words included idiosyncratic or erroneous transcriptions and/or repetitions of children's babbling. <!--kc: can do more thorough checks to report % and/or choose a less arbitrary cutoff-->], we calculated a rarity score as the negative log proportion of other-produced tokens in CHILDES (i.e., number of tokens for a given word/sum of all tokens in the full corpus), and then averaged for rarity scores for all target utterances. Utterances with more low-frequency words are considered more lexically rare.

### Syntactic level
Syntactic measures included both the utterance __length__ (in words) and __number of verb phrases__. The number of words per utterance was automatically extracted using the `childesr` package [@braginsky2021childesr]. The number of verb phrases per utterance was determined using `spaCy3`, an automatic syntactic parser [@honnibal2020spacy].

# Results

We ran individual mixed-effects binomial logistic regression models for each of seven linguisitc input predictors. Models included fixed effects of linguistic predictor, target child age, and their interaction as well as random intercepts for individual word pairs and speakers. For each target word token, form was coded as CDS (0) or ADS (1), so coefficient estimates should be interpreted as an indication of the likelihood that an utterance contains an ADS form. <!--kc: this wording needs work but leaving it for now--> All main effects of linguistic predictors and interactions with age are shown in Figure 3.

At the prosodic level, we found significant effects for two of three input predictors tested. Utterance-level __mean pitch__ was not predictive of form ($\beta$ = `r round(subset(input.models, Predictor == "Pitch mean")$Estimate, 3)`, _SE_ = `r round(subset(input.models, Predictor == "Pitch mean")$SE, 2)`, _t_ = `r round(subset(input.models, Predictor == "Pitch mean")$statistic, 2)`, _p_ `r ifelse(subset(input.models, Predictor == "Pitch mean")$p.value < .001, paste("< 0.001"), paste("=", round(subset(input.models, Predictor == "Pitch mean")$p.value, 3)))`) and did not significantly interact with age ($\beta$ = `r round(subset(input.models, Predictor == "Pitch mean:age")$Estimate, 2)`, _SE_ = `r round(subset(input.models, Predictor == "Pitch mean:age")$SE, 2)`, _t_ = `r round(subset(input.models, Predictor == "Pitch mean:age")$statistic, 2)`, _p_ `r ifelse(subset(input.models, Predictor == "Pitch mean:age")$p.value < .001, paste("< 0.001"), paste("=", round(subset(input.models, Predictor == "Pitch mean:age")$p.value, 3)))`). However, utterance-level __pitch range__ was a negative predictor of ADS form ($\beta$ = `r round(subset(input.models, Predictor == "Pitch range")$Estimate, 3)`, _SE_ = `r round(subset(input.models, Predictor == "Pitch range")$SE, 2)`, _t_ = `r round(subset(input.models, Predictor == "Pitch range")$statistic, 2)`, _p_ `r ifelse(subset(input.models, Predictor == "Pitch range")$p.value < .001, paste("< 0.001"), paste("=", round(subset(input.models, Predictor == "Pitch range")$p.value, 3)))`). That is, utterances with greater variability in pitch were more likely to contain CDS forms. This pitch variability effect increased in strength across developmental time, as evidenced by a positive interaction with age ($\beta$ = `r round(subset(input.models, Predictor == "Pitch range:age")$Estimate, 2)`, _SE_ = `r round(subset(input.models, Predictor == "Pitch range:age")$SE, 2)`, _t_ = `r round(subset(input.models, Predictor == "Pitch range:age")$statistic, 2)`, _p_ `r ifelse(subset(input.models, Predictor == "Pitch range:age")$p.value < .001, paste("< 0.001"), paste("=", round(subset(input.models, Predictor == "Pitch range:age")$p.value, 3)))`). __Speech rate__ (i.e., words produced per second) was a positive predictor of ADS form ($\beta$ = `r round(subset(input.models, Predictor == "Rate")$Estimate, 2)`, _SE_ = `r round(subset(input.models, Predictor == "Rate")$SE, 2)`, _t_ = `r round(subset(input.models, Predictor == "Rate")$statistic, 2)`, _p_ `r ifelse(subset(input.models, Predictor == "Rate")$p.value < .001, paste("< 0.001"), paste("=", round(subset(input.models, Predictor == "Rate")$p.value, 3)))`). Utterances spoken more quickly were more likely to contain ADS forms. This input predictor also negatively interacted with age ($\beta$ = `r round(subset(input.models, Predictor == "Rate:age")$Estimate, 2)`, _SE_ = `r round(subset(input.models, Predictor == "Rate:age")$SE, 2)`, _t_ = `r round(subset(input.models, Predictor == "Rate:age")$statistic, 2)`, _p_ `r ifelse(subset(input.models, Predictor == "Rate:age")$p.value < .001, paste("< 0.001"), paste("=", round(subset(input.models, Predictor == "Rate:age")$p.value, 3)))`), indicating a decreasing strength in predictive power across developmental time.

At the lexical level, we found significant effects for both input predictors tested. Utterances with higher levels of __lexical complexity__ ($\beta$ = `r round(subset(input.models, Predictor == "Complexity")$Estimate, 2)`, _SE_ = `r round(subset(input.models, Predictor == "Complexity")$SE, 2)`, _t_ = `r round(subset(input.models, Predictor == "Complexity")$statistic, 2)`, _p_ `r ifelse(subset(input.models, Predictor == "Complexity")$p.value < .001, paste("< 0.001"), paste("=", round(subset(input.models, Predictor == "Complexity")$p.value, 3)))`) and __lexical rarity__ ($\beta$ = `r round(subset(input.models, Predictor == "Rarity")$Estimate, 2)`, _SE_ = `r round(subset(input.models, Predictor == "Rarity")$SE, 2)`, _t_ = `r round(subset(input.models, Predictor == "Rarity")$statistic, 2)`, _p_ `r ifelse(subset(input.models, Predictor == "Rarity")$p.value < .001, paste("< 0.001"), paste("=", round(subset(input.models, Predictor == "Rarity")$p.value, 3)))`) were more likely to contain ADS forms. Lexical complexity did not interact with age ($\beta$ = `r round(subset(input.models, Predictor == "Complexity:age")$Estimate, 2)`, _SE_ = `r round(subset(input.models, Predictor == "Complexity:age")$SE, 2)`, _t_ = `r round(subset(input.models, Predictor == "Complexity:age")$statistic, 2)`, _p_ `r ifelse(subset(input.models, Predictor == "Complexity:age")$p.value < .001, paste("< 0.001"), paste("=", round(subset(input.models, Predictor == "Complexity:age")$p.value, 3)))`); whereas, lexical rarity negatively interacted with age such that there was a decreasing effect of this predictor with age ($\beta$ = `r round(subset(input.models, Predictor == "Rarity:age")$Estimate, 2)`, _SE_ = `r round(subset(input.models, Predictor == "Rarity:age")$SE, 2)`, _t_ = `r round(subset(input.models, Predictor == "Rarity:age")$statistic, 2)`, _p_ `r ifelse(subset(input.models, Predictor == "Rarity:age")$p.value < .001, paste("< 0.001"), paste("=", round(subset(input.models, Predictor == "Rarity:age")$p.value, 3)))`).

At the syntactic level, we found significant effects of __utterance length__ and __number of verb phrases__. Utterances with more words ($\beta$ = `r round(subset(input.models, Predictor == "Length")$Estimate, 2)`, _SE_ = `r round(subset(input.models, Predictor == "Length")$SE, 2)`, _t_ = `r round(subset(input.models, Predictor == "Length")$statistic, 2)`, _p_ `r ifelse(subset(input.models, Predictor == "Length")$p.value < .001, paste("< 0.001"), paste("=", round(subset(input.models, Predictor == "Length")$p.value, 3)))`) and more verb phrases ($\beta$ = `r round(subset(input.models, Predictor == "Verbs")$Estimate, 2)`, _SE_ = `r round(subset(input.models, Predictor == "Verbs")$SE, 2)`, _t_ = `r round(subset(input.models, Predictor == "Verbs")$statistic, 2)`, _p_ `r ifelse(subset(input.models, Predictor == "Verbs")$p.value < .001, paste("< 0.001"), paste("=", round(subset(input.models, Predictor == "Verbs")$p.value, 3)))`) were more likely to contain ADS forms. Moreover, both linguistic predictors negatively interacted with age (Length: $\beta$ = `r round(subset(input.models, Predictor == "Length:age")$Estimate, 2)`, _SE_ = `r round(subset(input.models, Predictor == "Length:age")$SE, 2)`, _t_ = `r round(subset(input.models, Predictor == "Length:age")$statistic, 2)`, _p_ `r ifelse(subset(input.models, Predictor == "Length:age")$p.value < .001, paste("< 0.001"), paste("=", round(subset(input.models, Predictor == "Length:age")$p.value, 3)))`; Verbs: $\beta$ = `r round(subset(input.models, Predictor == "Verbs:age")$Estimate, 2)`, _SE_ = `r round(subset(input.models, Predictor == "Verbs:age")$SE, 2)`, _t_ = `r round(subset(input.models, Predictor == "Verbs:age")$statistic, 2)`, _p_ `r ifelse(subset(input.models, Predictor == "Verbs:age")$p.value < .001, paste("< 0.001"), paste("=", round(subset(input.models, Predictor == "Verbs:age")$p.value, 3)))`), suggesting that the strength of these predictors decreases across developmental time.

# Discussion

<!--
# Acknowledgements
Research reported in this publication was supported by the Eunice Kennedy Shriver National Institute of Child Health & Human Development of the National Institutes of Health under Award Number P01HD040605; by a grant from the Successful Pathways from School to Work initiative of the University of Chicago, funded by the Hymen Milgrom Supporting Organization; by a grant from the Institute of Education Sciences, U.S. Department of Education, through grant R305A190467 to the University of Chicago; and by a grant from the Spencer Foundation. The content is solely the responsibility of the authors and does not necessarily represent the official views of the National Institutes of Health, the Institute of Education Sciences, or the U.S. Department of Education.
-->

# References

```{r}
# References will be generated automatically by Pandoc and included here.
# The following code is some latex to format the bibliography. Do not remove it.
```

\setlength{\parindent}{-0.1in} 
\setlength{\leftskip}{0.125in}
\noindent