---
title: "Developmental shifts in children's use of register-specific words"
bibliography: CogSci2022-register.bib
csl: apa6.csl
document-params: "10pt, letterpaper"

author-information: > 
    \author{{\large \bf Kennedy Casey} \\ University of Chicago \\ \texttt{kbcasey@uchicago.edu}
    \And {\large \bf Marisa Casillas} \\ University of Chicago \\ \texttt{mcasillas@uchicago.edu}}

abstract: >
  Child-directed language (CDL) features words such as _doggy_, _night-night_, and _tummy_ that are rarely used in adult-directed language (ADL). Characterisitcs of CDL word forms, such as diminutivization and reduplication, explain why they may be learned and produced earlier by children. However, it is not yet clear how or when children switch to using ADL equivalents---_dog_, _goodnight_, _stomach_. Through analysis of speech transcripts from CHILDES and the Language Development Project corpus, we show that children significantly increase their production of ADL word forms across age, with the average CDL-to-ADL transition point at 2.5 years. Many of the linguistic features that distinguish CDL vs. ADL registers (e.g., lexical and syntactic complexity) similarly differentiated the local speech contexts surrounding CDL vs. ADL word forms. Learners may therefore be able to capitalize on these cues to support their discovery of register along with context-appropriate CDL/ADL pair use.
  
keywords: >
    child-directed language; word production; linguistic input; social register; corpus analysis; developmental change
    
output: cogsci2016::cogsci_paper
final-submission: \cogscifinalcopy
---

```{r global_options, include = FALSE}
knitr::opts_chunk$set(fig.width = 3, fig.height = 3, fig.crop = F, 
                      fig.pos = "tb", fig.path = 'figs/',
                      echo = F, warning = F, cache = F, 
                      message = F, sanitize = T)
```

```{r, libraries}
library(png)
library(grid)
library(xtable)
library(kableExtra)
library(tidyverse)
library(ggpubr)
library(DescTools)
```

```{r, data prep}

# read in outputs from "get-freq.R"
metadata <- read_csv("../../data-prep/overall/metadata.csv")
freq <- read_csv("../../data-prep/overall/freq.csv")

cdl_forms <- read_csv("../../data-prep/overall/item-info.csv") %>%
    filter(form == "CDS") %>%
    pull(word)

adl_forms <- read_csv("../../data-prep/overall/item-info.csv") %>%
    filter(form == "ADS") %>%
    pull(word)

# preset colors for linguistic predictors
input_colors <- c("Prosodic" = "#1C9E78", 
                  "Lexical" = "#D95F06", 
                  "Syntactic" = "#7570B4")

# get counts of child-produced and other-produced utterances
child.utts <- 64852 #calculated in "shift-timing.R"

other.utts <- read_csv("../../data/childes-input.csv") %>%
  select(id) %>%
  distinct() %>%
  nrow()

caregiver.utts <- read_csv("../../data/childes-input.csv") %>%
  filter(speaker_code %in% c("MOT", "FAT")) %>%
  select(id) %>%
  distinct() %>%
  nrow()
```

# Introduction
Across their first few years of life, children amass an impressive expressive vocabulary that includes hundreds if not thousands of words [@fenson1994variability; @mayor2011statistical]. Word production typically begins around age one, followed by a vocabulary 'explosion' or 'spurt' during toddlerhood [@ganger2004reexamining; see also @mcmurray2007defusing] and continued, measurable growth in vocabulary size thereafter [@rice2015predicting].

## Predictors of early word comprehension and production
Which words do children understand and say first? Cross-linguistic predictors of early age of acquisition include overall frequency, frequency of occurrence in short or isolated utterances, and concreteness [@braginsky2019consistency]. Words that adults rate as more associated with infancy (e.g., _bottle_ and _bib_) are typically understood earlier [@perry2015iconicity], and shorter words with fewer phonemes are also produced earlier [@braginsky2019consistency].

Additional features of word forms, such as iconicity [@laing2017salient], diminutivization [@kempe2005diminutives], and reduplication [@ota2018choo], have been shown to support children's early learning. Words with these features (e.g., _doggy_ and _choo-choo_) are typically overrepresented in children's early productive vocabularies [@perry2018iconicity; @frank2017wordbank] and serve as a hallmark of child-directed language (CDL) across many communities [@ferguson1964baby].

In sum, children's productive vocabulary gets off the ground with many words that are particularly relevant for them <!--kc: wording needs some work here-->. These include words for the key people, routines, and objects in their early lives [e.g., @tardif2008baby; @frank2021wordbank], along with words that appear specifically tailored to their level of linguisitc maturity (i.e., CDL-specific words). However, at some point (likely gradually) children shift away from these baby-centric early words. This transition is perhaps most marked for common child language word forms that have direct substitutes in adult language (e.g., _doggy/dog_, _tummy/stomach_, _bunny/rabbit_; hereafter, 'CDL/ADL word pairs') as opposed to words that become less relevant over the years (e.g., _diaper_, _pacifier_, _peekaboo_). Note, though, that these CDL-specific words do not fully disappear---children themselves come to use these lexical items (along with other CDL features) when talking to infants and younger children (REFS).

## Comprehension and production of language varieties
Classically, we might expect that hearing multiple labels for the same referent (i.e., both CDL and ADL words) would be a problem for early word learning---particularly when variants have little to no overlap in phonological form (e.g., _tummy/stomach_, _bunny/rabbit_).<!--kc: maybe a point here about this being the motivation to look at CDL words in most past research--> Indeed, children often apply a mutual exclusivity (ME) heuristic when learning new words; they assume that new labels refer to new items rather than interpreting them synonyms for words they already know [@markman1988children; see @lewis2020role, for a recent meta-analysis]. However, children seem to contend with multiple CDL/ADL variants without issue. 

One possible way to explain the relative ease of learning both labels is to consider the social context of CDL vs. ADL use. While labeling an animal as _doggy_ vs. _dog_ may not communicate anything distinct about the referent itself, the production of one form vs. the other may indicate something meaningful about _who_ is hearing or producing the label (i.e., CDL and children's own productions are more likely to feature talk about _doggies_, while we would expect much more frequent talk about _dogs_ in ADL). That is, differences in register could serve to 'explain away' the otherwise problematic redundancy of multiple labels [@clark1987principle].

This account seems theoretically plausible given evidence that children become increasingly aware of the fact that language style is modulated by a variety of social factors, including the identities of speakers [e.g., from social groups: @liberman2017preverbal] along with their addressees [e.g., young children vs. adults: @soley2020infants]. Moreover, we know that children are exposed to many forms of socially meaningful variation in their linguistic input [see @johnson2020developmental, for a recent review] -- not just CDL vs. ADL words and registers [@bunceURcross; @loukatou2021child] but many different speakers [@bulgarelli2021talker; @rost2009speaker], accents [@buckler2017input; @potter2017exposure], dialects [@durrant2015monodialectal; @edwards2014dialect], and languages [@kremin2020code]. Rather than viewing this variation as noise or a complication for learning, accumulating evidence suggests that experience with variable input affords children greater ability to detect, learn from, and apply social meaning to such variation. For instance, children's tendency to show the ME effect is modulated by their experience with multiple languages [and thus multiple labels: @byers2009monolingual; @houston2010language] as well as the social conditions under which multiple labels are introduced [e.g., by speakers of familiar or unfamiliar race: @weatherhead2021putting].

Returning to the 'issue' of multiple CDL/ADL variants, we hypothesize that children may be able to make sense of this lexical variation by beginning to associate different forms with different modes of use (i.e., incipient representations of register). To test this idea, we first need to establish (a) when children begin to shift away from producing CDL-specific words, and (b) how children may be able to use bottom-up linguistic input cues to associate lexical variants with their appropriate registers (i.e., CDL vs. ADL).

## Current investigation
We examine a core subset of 15 CDL-specific words in English (e.g., _doggy_, _night-night_, _tummy_) that are prevalent in children’s early vocabularies but are eventually ‘replaced’ by ADL words---_dog_, _goodnight_, _stomach_. In Study 1, we analyze `r format(child.utts, big.mark = ",", trim = TRUE)` utterances of spontaneous speech from children up to 7 years of age to establish when ADL forms become more dominant in children’s own productions. That is, when do children 'switch' from using CDL to ADL forms? 

After establishing that the average age of ‘switchover’ occurs around 2.5 years<!--kc: i wonder if we shouldn't make the 50/50 point so marked-->, we next explored the features children's input that could support this developmental shift. We investigated the extent to which CDL and ADL words are used in distinct linguistic contexts. Further processing of `r format(other.utts, big.mark = ",", trim = TRUE)` adult utterances revealed that CDL and ADL variants co-occur with reliably different patterns of prosodic, lexical, and syntactic information---cues that likely help learners associate them with different modes of use, or emerging representations of register.

Together, these studies push us to consider children’s vocabulary development not as a simple accumulation of words or numeric increase in vocabulary size but rather a deepening and restructuring of the lexicon with growing linguistic and social maturity. The word _dog_ does not entirely replace _doggy_---rather, the two forms become reserved for use with different addressees.

# Study 1: When do children shift from CDL to ADL forms?
We tracked children's use of 15 CDL/ADL word pairs (Table 1) from early infancy up to age seven. Since CDL forms rarely appear in ADL, we predicted that children would shift away from production of these CDL-specific forms with increasing age. That is, we expected to see 'replacement' of CDL forms with ADL forms in children's own speech across time. 

# Method

## Corpora

We analyzed `r metadata$n_transcripts` transcripts in the North American English collection of the Child Language Data Exchange System (CHILDES) database [@macwhinney2000childes] for children up to 7 years of age. The included transcripts were drawn from `r metadata$n_corpora` individual corpora and featured `r metadata$n_kids` children (age range = `r metadata$min_age`--`r metadata$max_age` months, _M_ = `r metadata$mean_age` months). Child production data from the Language Development Project longitudinal corpus were also analyzed [see @huttenlocher2010sources; @rowe2008child, for further details regarding participating families, recording procedures, and transcription]. These included 622 transcripts from 59 English-learning children recorded every 4 months for approximately 90 minutes from age 14 to 58 months.

```{r freq_table, results = "asis"}
cdl_freq <- freq %>%
    filter(word %in% cdl_forms)

adl_freq <- freq %>%
    filter(word %in% adl_forms)

freq <- bind_cols(cdl_freq, adl_freq) %>%
  mutate(pair = paste0(word...1, '/', word...4)) %>%
  select(pair, child...2, other...3, child...5, other...6) %>%
  mutate(child...2 = paste0(format(child...2, big.mark = ",", trim = TRUE)), 
         child...5 = paste0(format(child...5, big.mark = ",", trim = TRUE)), 
         other...3 = paste0(format(other...3, big.mark = ",", trim = TRUE)), 
         other...6 = paste0(format(other...6, big.mark = ",", trim = TRUE)))
  
colnames(freq) <- c("Pair", "Child", "Other", "Child", "Other")

kable(freq, format = "latex", 
      table.envir = "table", 
      row.names = FALSE, 
      label = "tab:tab1",
      booktabs = TRUE,
      position = "h",
      align = "lcccc",
      linesep = "",
      caption = "CHILDES frequency for 15 CDL/ADL word pairs. Child-produced counts include tokens produced only by the target child.") %>%
  kable_styling(latex_options = "scale_down") %>%
  add_header_above(c(" " = 1, "CDL tokens by speaker" = 2, "ADL tokens by speaker" = 2), bold = TRUE) %>%
  row_spec(0, bold = TRUE) %>%
  column_spec(1, italic = TRUE)
```

## Target words
Fifteen CDL/ADL word pairs (30 total target words) were selected based on two criteria: the appearance of at least one form on the MacArthur-Bates Communicative Development Inventory [CDI, @fenson1994variability], and sufficient frequency of occurrence in CHILDES (at least 100 child-produced tokens and 100 other-produced tokens per form). Pairs were also selected based on researcher intuition to ensure that the same object, animal, routine, or body part could be reasonably labeled with either form by young children^[While onomatopoeic words can be used in a similar manner to the CDL-specific words in our test set (e.g., _choo-choo_ serving as a CDL-specific label for _train_, or _quack-quack_ for _duck_), these items were not included here because they are primarily used as sound effects rather than labels for objects or animals [@skarabela2018train]. Thus, the polysemous nature of iconic word usage does not provide as clear of a test of 'replacement' of CDL forms with ADL forms over time.] Across all transcripts, `r format(child.utts, big.mark = ",", trim = TRUE)` child-produced utterances contained at least one target word and were included in our analysis.

# Results
```{r}
# read in lmer outputs generated in shift-timing.R 
overall.shift <- read_csv("../../analysis/model-outputs/shift-timing.csv")

# get count of items with sig increase in ADL with age
bypair.shift <- read_csv("../../analysis/model-outputs/bypair-shift-timing.csv") %>%
  filter(p.value < 0.05) %>%
  nrow()
```
We asked when CDL forms are 'replaced' by ADL forms in children's own speech. Using the `lme4` package [version 1.1.27.1, @bates2015fitting] in R [version 4.1.0, @r2021], we fit a mixed-effects binomial logistic regression model predicting children's production of CDL vs. ADL forms, with target child age (in months, scaled) as a single fixed effect. Random slopes and intercepts for word pairs were also included. For each target word token, the form was coded as either 0 (CDL) or 1 (ADL). Thus, the model captures, for each age, the relative proportion of CDL vs. ADL forms in children's own speech.

Children's production of ADL forms increased with age ($\beta$ = `r format(round(overall.shift$estimate, 2), nsmall = 2)`, _SE_ = `r format(round(overall.shift$std.error, 2), nsmall = 2)`, _t_ = `r format(round(overall.shift$statistic, 2), nsmall = 2)`, _p_ `r ifelse(overall.shift$p.value < .001, paste("< 0.001"), paste("=", format(round(overall.shift$p.value, 3), 3)))`). The average CDL-to-ADL transition point (i.e., the point at which ADL forms were produced >50% of the time) was between 24 and 36 months of age (Figure 1). 

The trend of increasing ADL form production was significant for `r bypair.shift` of 15 word pairs, but the exact trajectory of shift varied greatly across pairs (Figure 2). In some cases, CDL forms were 'replaced' by ADL forms early on (e.g., _doggy/dog_ and _kitty/cat_). For other pairs, the age 'switchover' was much later (e.g., _tummy/stomach_ and _potty/bathroom_) or was not observed in our sample because ADL forms were always produced >50% of the time (e.g., _duckie/duck_ and _blankie/blanket_).

```{r shift-timing-fig, fig.env = "figure", fig.pos = "h", fig.align = "left", fig.width = 3.4, fig.height = 3.4, fig.cap = "Model-predicted increase in production of ADL forms across age, with shaded standard error region. Gray lines depict individual word-pair trajectories."}

shift.timing <- png::readPNG("../figs/shift-timing.png")
grid::grid.raster(shift.timing)
```

```{r shift-timing-bypair-fig, fig.env = "figure*", fig.pos = "!ht", fig.align = "center", fig.width = 11, fig.height = 4, fig.cap = "Inidividual word-pair trajectories for increasing production of ADL forms (blue) and descreasing production of CDL forms (red) with age. Points indicate proportions for each 1-month age bin."}

bypair.shift.timing <- png::readPNG("../figs/bypair-shift-timing.png")
grid::grid.raster(bypair.shift.timing)
```

# Discussion

# Study 2: What linguisitc information in children's input supports their shift from CDL to ADL forms?
We next explored children's input (i.e., other-produced speech), asking whether the linguistic features that differentiate CDL vs. ADL at the register level also differentiate the local speech contexts surrounding CDL vs. ADL forms. In other words, can form be predicted on the basis of individual utterance-level prosodic, lexical, or syntactic cues?

```{r ling-predictors-fig, fig.env = "figure*", fig.pos = "h", fig.align = "center", fig.width = 7,  set.cap.width = TRUE, num.cols.cap = 2, fig.cap = "Coefficient estimates for linguistic predictors of form. Positive main effects indicate that utterances are more likely to contain ADL forms when they have higher values for the predictor (e.g., faster speech rates). Positive age interactions indicate an increasing effect of the predictor with age. Error bars depict standard errors of the coefficient estimates, and filled circles represent significant effects (\\textit{p} $<$ 0.05)."}

path <- "../../analysis/model-outputs/input/"
filenames <- list.files(path, "*.csv")
files <- lapply(paste0(path, filenames), read_csv)
input.models <- bind_rows(files, .id = "source") %>%
  rename(Predictor = term, 
         Estimate = estimate, 
         SE = std.error) %>%
  group_by(source) %>%
  mutate(order = row_number(), 
         type = case_when(
           order == 1 ~ "intercept", 
           order == 2 ~ "main_effect", 
           order == 3 ~ "age", 
           order == 4 ~ "interaction"), 
         Sig = ifelse(p.value < .05, "sig", "not_sig"), 
         Predictor = str_to_sentence(str_replace(str_remove_all(Predictor, "_scaled"), "_", " "), "first"), 
         level = ifelse(str_detect(Predictor, "Length|Verbs"), "Syntactic", 
                        ifelse(str_detect(Predictor, "Complexity|Rarity"), "Lexical", "Prosodic"))) %>%
  mutate(Predictor = factor(Predictor, levels = c("Verbs", "Verbs:age", "Length", "Length:age", "Rarity", "Rarity:age", "Complexity", "Complexity:age", "Rate", "Rate:age", "Pitch range", "Pitch range:age", "Pitch mean", "Pitch mean:age")))

main.effects <- ggplot(filter(input.models, type == "main_effect"), 
       aes(x = Estimate, y = Predictor, color = level)) +
  geom_vline(xintercept = 0, size = 0.75, linetype = "dotted", 
             color = "gray", alpha = 0.5) +
  geom_errorbar(aes(xmin = Estimate - SE, xmax = Estimate + SE), 
                width = 0, size = 0.75) + 
  geom_point(aes(shape = Sig), size = 1.5) +
  scale_shape_manual(values = c(21, 19)) +
  scale_color_manual(values = input_colors) +
  scale_x_continuous(limits = c(-0.5, 0.5)) +
  labs(title = "Main Effect", x = "Coefficient Estimate") +
  theme_test(base_size = 10) +
  theme(legend.position = "none", plot.title = element_text(hjust = 0.5))
  
interactions <- ggplot(filter(input.models, type == "interaction"), 
       aes(x = Estimate, y = Predictor, color = level)) +
  geom_vline(xintercept = 0, size = 0.5, linetype = "dotted", 
             color = "gray", alpha = 0.75) +
  geom_errorbar(aes(xmin = Estimate - SE, xmax = Estimate + SE), 
                width = 0, size = 0.75) + 
  geom_point(aes(shape = Sig), size = 1.5) +
  scale_shape_manual(values = c(21, 19), guide = "none") +
  scale_color_manual(values = input_colors) +
  scale_x_continuous(limits = c(-0.5, 0.5)) +
  labs(title = "Interaction with Age", x = "Coefficient Estimate") +
  theme_test(base_size = 10) +
  theme(legend.position = "right", legend.title = element_blank(), 
        plot.title = element_text(hjust = 0.5), 
        axis.text.y = element_blank(), axis.title.y = element_blank(), 
        axis.ticks.y = element_blank())

model.fig <- ggarrange(main.effects, interactions, widths = c(3, 3.2))
ggsave("../figs/model-outputs.png", model.fig, height = 3, width = 5, dpi = 300)

img <- png::readPNG("../figs/model-outputs.png")
grid::grid.raster(img)
```

# Method

## Corpora
We analyzed `r format(other.utts, big.mark = ",", trim = TRUE)` other-produced utterances (i.e., utterances not produced by the target child) in the same CHILDES transcripts from Study 1. The majority were utterances from primary caregivers (_n_ = `r format(caregiver.utts, big.mark = ",", trim = TRUE)`, or `r format(round(caregiver.utts/other.utts, 3)*100, nsmall = 1)`%).

## Linguistic input predictors

All input analyses were conducted over individual utterances containing one of the 30 target words from Study 1. We quantified prosodic, lexical, and syntactic information to describe each utterance.

### Prosodic level

We measured three types of prosodic information: __mean pitch__ (Hz), __pitch range__ (Hz), and __speech rate__ (words per second). These measures were calculated over all timestamped utterances in CHILDES (`r metadata$other_timestamped`% of other-produced utterances). Utterances shorter than 58 ms were excluded from analysis. This lower bound was set by identifying the the shortest possible duration of an utterance containing at least one word in four manually annotated North American English corpora in HomeBank [@homebank; @bergelsoncorpus; @soderstromcorpus; @vandamcorpus; @warlaumontcorpus] <!--kc2mc: are these the right refs?-->. Pitch information was extracted using Praat software [@boersma2016praat].

### Lexical level

We measured two types of lexical information: complexity and rarity. __Lexical complexity__ was defined as the negative log proportion of known words in each utterance [consistent with @foushee2016lexical; @kidd2012goldilocks]. A word was considered 'known' if the age of acquisition (AoA) estimate [@fenson1994variability; @frank2017wordbank] was less than or equal to the age of the target child when they heard the utterance. Utterances with proportionally fewer known words are more lexically complex. __Lexical rarity__ was determined based on overall frequency in CHILDES. For all words with at least 10 tokens^[Manual checks revealed that many of the lowest-frequency words included idiosyncratic or erroneous transcriptions, so we excluded words with <10 tokens from our estimates of lexical rarity to reduce noise in this measure.], we calculated a rarity score as the negative log proportion of other-produced tokens in CHILDES (i.e., number of tokens for a given word/sum of all tokens in the full corpus), and then averaged for rarity scores for all target utterances. Utterances with more low-frequency words are considered more lexically rare.

### Syntactic level

Syntactic measures included both the utterance __length__ (in words) and __number of verb phrases__. The number of words per utterance was automatically extracted using the `childesr` package [@braginsky2021childesr]. The number of verb phrases per utterance was determined using `spaCy3`, an automatic syntactic parser [@honnibal2020spacy].

# Results

We ran individual mixed-effects binomial logistic regression models for each of seven linguisitc input predictors. Models included fixed effects of linguistic predictor, target child age, and their interaction as well as random intercepts for individual word pairs and speakers. For each target word token, form was coded as CDL (0) or ADL (1), so coefficient estimates should be interpreted as an indication of the likelihood that an utterance contains an ADL form. <!--kc: this wording needs work but leaving it for now--> All main effects of linguistic predictors and interactions with age are shown in Figure 3.

At the prosodic level, we found significant effects for two of three input predictors tested. Utterance-level __pitch range__ was not predictive of form ($\beta$ = `r round(subset(input.models, Predictor == "Pitch range")$Estimate, 3)`, _SE_ = `r round(subset(input.models, Predictor == "Pitch range")$SE, 2)`, _t_ = `r round(subset(input.models, Predictor == "Pitch range")$statistic, 2)`, _p_ `r ifelse(subset(input.models, Predictor == "Pitch range")$p.value < .001, paste("< 0.001"), paste("=", round(subset(input.models, Predictor == "Pitch range")$p.value, 3)))`) and did not significantly interact with age ($\beta$ = `r round(subset(input.models, Predictor == "Pitch range:age")$Estimate, 2)`, _SE_ = `r round(subset(input.models, Predictor == "Pitch range:age")$SE, 2)`, _t_ = `r round(subset(input.models, Predictor == "Pitch range:age")$statistic, 2)`, _p_ `r ifelse(subset(input.models, Predictor == "Pitch range:age")$p.value < .001, paste("< 0.001"), paste("=", round(subset(input.models, Predictor == "Pitch range:age")$p.value, 3)))`). However, utterance-level __mean pitch__ was a negative predictor of ADL form ($\beta$ = `r round(subset(input.models, Predictor == "Pitch mean")$Estimate, 3)`, _SE_ = `r round(subset(input.models, Predictor == "Pitch mean")$SE, 2)`, _t_ = `r round(subset(input.models, Predictor == "Pitch mean")$statistic, 2)`, _p_ `r ifelse(subset(input.models, Predictor == "Pitch mean")$p.value < .001, paste("< 0.001"), paste("=", round(subset(input.models, Predictor == "Pitch mean")$p.value, 3)))`). That is, utterances with higher overall mean pitch were more likely to contain CDL forms, with no significant interaction with age ($\beta$ = `r round(subset(input.models, Predictor == "Pitch mean:age")$Estimate, 2)`, _SE_ = `r round(subset(input.models, Predictor == "Pitch mean:age")$SE, 2)`, _t_ = `r round(subset(input.models, Predictor == "Pitch mean:age")$statistic, 2)`, _p_ `r ifelse(subset(input.models, Predictor == "Pitch mean:age")$p.value < .001, paste("< 0.001"), paste("=", round(subset(input.models, Predictor == "Pitch mean:age")$p.value, 3)))`). __Speech rate__ (i.e., words produced per second) was a positive predictor of ADL form ($\beta$ = `r round(subset(input.models, Predictor == "Rate")$Estimate, 2)`, _SE_ = `r round(subset(input.models, Predictor == "Rate")$SE, 2)`, _t_ = `r round(subset(input.models, Predictor == "Rate")$statistic, 2)`, _p_ `r ifelse(subset(input.models, Predictor == "Rate")$p.value < .001, paste("< 0.001"), paste("=", round(subset(input.models, Predictor == "Rate")$p.value, 3)))`). Utterances spoken more quickly were more likely to contain ADL forms. This input predictor also negatively interacted with age ($\beta$ = `r round(subset(input.models, Predictor == "Rate:age")$Estimate, 2)`, _SE_ = `r round(subset(input.models, Predictor == "Rate:age")$SE, 2)`, _t_ = `r round(subset(input.models, Predictor == "Rate:age")$statistic, 2)`, _p_ `r ifelse(subset(input.models, Predictor == "Rate:age")$p.value < .001, paste("< 0.001"), paste("=", round(subset(input.models, Predictor == "Rate:age")$p.value, 3)))`), indicating a decreasing strength in predictive power across developmental time.

At the lexical level, we found significant effects for both input predictors tested. Utterances with higher levels of __lexical complexity__ ($\beta$ = `r round(subset(input.models, Predictor == "Complexity")$Estimate, 2)`, _SE_ = `r round(subset(input.models, Predictor == "Complexity")$SE, 2)`, _t_ = `r round(subset(input.models, Predictor == "Complexity")$statistic, 2)`, _p_ `r ifelse(subset(input.models, Predictor == "Complexity")$p.value < .001, paste("< 0.001"), paste("=", round(subset(input.models, Predictor == "Complexity")$p.value, 3)))`) and __lexical rarity__ ($\beta$ = `r round(subset(input.models, Predictor == "Rarity")$Estimate, 2)`, _SE_ = `r round(subset(input.models, Predictor == "Rarity")$SE, 2)`, _t_ = `r round(subset(input.models, Predictor == "Rarity")$statistic, 2)`, _p_ `r ifelse(subset(input.models, Predictor == "Rarity")$p.value < .001, paste("< 0.001"), paste("=", round(subset(input.models, Predictor == "Rarity")$p.value, 3)))`) were more likely to contain ADL forms. Lexical complexity did not interact with age ($\beta$ = `r format(round(subset(input.models, Predictor == "Complexity:age")$Estimate, 2), nsmall = 2)`, _SE_ = `r round(subset(input.models, Predictor == "Complexity:age")$SE, 2)`, _t_ = `r round(subset(input.models, Predictor == "Complexity:age")$statistic, 2)`, _p_ `r ifelse(subset(input.models, Predictor == "Complexity:age")$p.value < .001, paste("< 0.001"), paste("=", round(subset(input.models, Predictor == "Complexity:age")$p.value, 3)))`); whereas, lexical rarity negatively interacted with age such that there was a decreasing effect of this predictor with age ($\beta$ = `r round(subset(input.models, Predictor == "Rarity:age")$Estimate, 2)`, _SE_ = `r round(subset(input.models, Predictor == "Rarity:age")$SE, 2)`, _t_ = `r round(subset(input.models, Predictor == "Rarity:age")$statistic, 2)`, _p_ `r ifelse(subset(input.models, Predictor == "Rarity:age")$p.value < .001, paste("< 0.001"), paste("=", round(subset(input.models, Predictor == "Rarity:age")$p.value, 3)))`).

At the syntactic level, we found significant effects of __utterance length__ and __number of verb phrases__. Utterances with more words ($\beta$ = `r round(subset(input.models, Predictor == "Length")$Estimate, 2)`, _SE_ = `r round(subset(input.models, Predictor == "Length")$SE, 2)`, _t_ = `r format(round(subset(input.models, Predictor == "Length")$statistic, 2), nsmall = 2)`, _p_ `r ifelse(subset(input.models, Predictor == "Length")$p.value < .001, paste("< 0.001"), paste("=", round(subset(input.models, Predictor == "Length")$p.value, 3)))`) and more verb phrases ($\beta$ = `r round(subset(input.models, Predictor == "Verbs")$Estimate, 2)`, _SE_ = `r round(subset(input.models, Predictor == "Verbs")$SE, 2)`, _t_ = `r round(subset(input.models, Predictor == "Verbs")$statistic, 2)`, _p_ `r ifelse(subset(input.models, Predictor == "Verbs")$p.value < .001, paste("< 0.001"), paste("=", round(subset(input.models, Predictor == "Verbs")$p.value, 3)))`) were more likely to contain ADL forms. Moreover, both linguistic predictors negatively interacted with age (Length: $\beta$ = `r round(subset(input.models, Predictor == "Length:age")$Estimate, 2)`, _SE_ = `r round(subset(input.models, Predictor == "Length:age")$SE, 2)`, _t_ = `r round(subset(input.models, Predictor == "Length:age")$statistic, 2)`, _p_ `r ifelse(subset(input.models, Predictor == "Length:age")$p.value < .001, paste("< 0.001"), paste("=", round(subset(input.models, Predictor == "Length:age")$p.value, 3)))`; Verbs: $\beta$ = `r round(subset(input.models, Predictor == "Verbs:age")$Estimate, 2)`, _SE_ = `r round(subset(input.models, Predictor == "Verbs:age")$SE, 2)`, _t_ = `r format(round(subset(input.models, Predictor == "Verbs:age")$statistic, 2), nsmall = 2)`, _p_ `r ifelse(subset(input.models, Predictor == "Verbs:age")$p.value < .001, paste("< 0.001"), paste("=", round(subset(input.models, Predictor == "Verbs:age")$p.value, 3)))`), suggesting that the strength of these predictors decreases across developmental time.

## Possible section talking about how some of these patterns are mirrored in children's own speech

# Discussion

# General Discussion

## Distinguishing CDL vs. ADL registers and words: Which cues matter?
Para about how we don't yet know which (if any of these) linguistic cues are actually exploited by learners [@potter2019infants as a good ref]. However, the current work sets us up to be able to test some of these ideas experimentally. Additional point that CDL and ADL registers are differentiated by more than just linguistic cues, so we can think about whether the same happens for CDL vs. ADL forms.

## Developing linguistic and social knowledge in parallel
Children's linguistic knowledge builds around and together with their social knowledge.

Lexical variants of CDL vs. ADL registers as one example of socially meaningful linguistic variation

Para on ties to other sociolinguistic phenomena, like dialects, accents, and other types of register (e.g., pedagogical, narrative, etc.). Learning happening at multiple levels -- not just words but associations between words and surrounding context (linguistic, social, etc.).

Word learning is clearly not a process of one-to-one mapping. Here, we focus on the multiple labels for the same referent, but there's also the opposite problem: one label for many different referents [@meylan2021quantifying; @caseyURmoving].

## More than vocabulary size: Understanding words _and_ using them in contextually-appropriate ways
Maybe some critique of typical caregiver-report measures, like CDIs, that gloss over lexical variation. Standardization makes sense for some things, but we lose the opportunity to capture deepening linguistic knowledge 

<!--
# Acknowledgements
We are grateful to the members of the University of Chicago Chatter Lab and Northwestern University Child Language Lab for valuable discussion and feedback on this work. Research reported in this publication was supported by the Eunice Kennedy Shriver National Institute of Child Health & Human Development of the National Institutes of Health under Award Number P01HD040605; by a grant from the Successful Pathways from School to Work initiative of the University of Chicago, funded by the Hymen Milgrom Supporting Organization; by a grant from the Institute of Education Sciences, U.S. Department of Education, through grant R305A190467 to the University of Chicago; and by a grant from the Spencer Foundation. The content is solely the responsibility of the authors and does not necessarily represent the official views of the National Institutes of Health, the Institute of Education Sciences, or the U.S. Department of Education.
-->

# References

```{r}
# References will be generated automatically by Pandoc and included here.
# The following code is some latex to format the bibliography. Do not remove it.
```

\setlength{\parindent}{-0.1in} 
\setlength{\leftskip}{0.125in}
\noindent