---
title: "Children's shift from CDS to ADS vocabulary across early childhood"
bibliography: CogSci2022.bib
csl: apa6.csl
document-params: "10pt, letterpaper"

author-information: > 
    \author{{\large \bf Kennedy Casey (kbcasey@uchicago.edu)} 
    \\{\large \bf Marisa Casillas (mcasillas@uchicago.edu)} \\Comparative Human Development, University of Chicago \\ Chicago, IL 60637 USA}

abstract: >
  abstract text here. abstract text here. abstract text here. abstract text here. abstract text here. abstract text here.
  abstract text here. abstract text here. abstract text here. abstract text here. abstract text here. abstract text here.
  abstract text here. abstract text here. abstract text here. abstract text here. abstract text here. abstract text here.
  abstract text here. abstract text here. abstract text here. abstract text here. abstract text here. abstract text here.
  abstract text here. abstract text here. abstract text here. abstract text here. abstract text here. abstract text here.
  abstract text here. abstract text here. abstract text here. abstract text here. abstract text here. abstract text here.
  abstract text here. abstract text here. abstract text here. abstract text here. abstract text here. abstract text here.
  abstract text here. abstract text here. abstract text here. abstract text here. abstract text here. abstract text here.
    
keywords: >
    child-directed speech; word production; linguistic input; social register; corpus analysis; developmental change
    
output: cogsci2016::cogsci_paper
#final-submission: \cogscifinalcopy
---

```{r global_options, include=FALSE}
knitr::opts_chunk$set(fig.width = 3, fig.height = 3, fig.crop = F, 
                      fig.pos = "tb", fig.path = 'figs/',
                      echo = F, warning = F, cache = F, 
                      message = F, sanitize = T)
```

```{r, libraries}
library(png)
library(grid)
library(xtable)
library(kableExtra)
library(tidyverse)

# source(data-prep/overall/get_freq.R)
metadata <- read_csv("../../data-prep/overall/metadata.csv")
freq <- read_csv("../../data-prep/overall/freq.csv")

cds_forms <- read_csv("../../data-prep/overall/item-info.csv") %>%
    filter(form == "CDS") %>%
    pull(word)

ads_forms <- read_csv("../../data-prep/overall/item-info.csv") %>%
    filter(form == "ADS") %>%
    pull(word)
```

# Introduction

# Methods

## Corpus and item selection

We analyzed `r metadata$n_transcripts` transcripts in the North American English collection of the Child Language Data Exchange System (CHILDES) database [@macwhinney2000childes]. The included transcripts were drawn from `r metadata$n_corpora` individual corpora and featured `r metadata$n_kids` children up to 7 years of age (range = `r metadata$min_age`--`r metadata$max_age` months, *M* = `r metadata$mean_age` months).

15 CDS/ADS word pairs were selected as test items based on their appearance on the MacArthur-Bates Communicative Development Inventory [@fenson1994variability] and their frequency of occurrence in CHILDES (at least 100 child-produced tokens and 100 other-produced tokens; see Table \ref{tab:tab1}).

Cases where the same object, animal, or routine could be reasonably labeled with either form in typical communicative interactions with young children.

```{r freq_table, results = "asis"}
cds_freq <- freq %>%
    filter(word %in% cds_forms)

ads_freq <- freq %>%
    filter(word %in% ads_forms)

freq <- bind_cols(cds_freq, ads_freq) %>%
    mutate(pair = paste0(word...1, '/', word...4)) %>%
    select(pair, child...2, other...3, child...5, other...6) 
colnames(freq) <- c("Pair", "Child", "Other", "Child", "Other")

freq_table <- xtable::xtable(freq, digits = 0,
                             include.rownames = FALSE, 
                             label = "tab:tab1",
                             caption = "CHILDES frequency for 15 CDS/ADS word pairs. Child-produced counts 
                             include tokens produced only by the target child. All other 
                             speakers' productions are included in the other-produced counts.") %>%
    xtable2kable(booktabs = TRUE, escape = FALSE) %>%
    kable_styling(font_size = 6) %>%
    add_header_above(c(" " = 2, "CDS tokens by speaker" = 2, "ADS tokens by speaker" = 2), bold = TRUE) %>%
    row_spec(0, bold = TRUE) %>%
    column_spec(2, italic = TRUE)

print(freq_table, include.rownames = FALSE,
      type = "latex", comment = FALSE, table.placement = "h!")
```

## Linguistic predictors

All analyses were conducted over individual utterances. We quantified prosodic, lexical and syntactic information to describe each utterance containing one of the 30 target words.

### Prosodic level

For all timestamped utterances (`r metadata$child_timestamped`% of child-produced and `r metadata$other_timestamped`% of other-produced utterances), mean pitch and pitch range were extracted using Praat software [@boersma2016praat]. Speech rate was calculated as the number of words produced per second. Utterances shorter than 58 ms were excluded from analysis. This lower bound was set by identifying the the shortest possible duration of an utterance containing at least one word in four manually annotated North American English corpora in HomeBank [@homebank; @bergelsoncorpus; @soderstromcorpus; @vandamcorpus; @warlaumontcorpus; see also @bergelson2019north].

### Lexical level

Typically, measures of lexical complexity are calculated over at least minutes or hours of transcribed speech rather than individual utterances. Here, we aimed to estimate local lexical complexity in two ways.

First, we calculated the negative log proportion of known words in each utterance [consistent with @foushee2016lexical; @kidd2012goldilocks]. A word was considered 'known' if the age of acquisition estimate (AoA) [@fenson1994variability, @frank2017wordbank] was less than or equal to the age of the target child when they heard or produced the utterance. Utterances with proportionally fewer known words are *more* lexically complex.

To account for more individual variation in word production, we included a measure

### Syntactic level

Syntactic measures included both the number of verb phrases and length (in words) of each utterance. Words were chosen over morphemes in the length analysis because we identified systematic errors in automatic morpheme counts. Manual checking of 5% of target utterances (i.e., XXXX utterances) revealed an error rate of XX%.

# Results

## Measuring production: When do children produce CDS vs. ADS forms?

## Characterizing the input: In what linguistic contexts do children hear CDS vs. ADS forms?

We used mixed-effects binomial logistic regression models to predict the appearance of CDS vs. ADS forms in given utterance on the basis of target child's age, several linguistic properties of the utterance, and interactions between each property and age. Models included random slopes and intercepts for individual word pairs and speakers and were fitted to all utterance data from speakers other than the target child.

## Modeling learning: What linguistic information can be used to distinguish CDS vs. ADS contexts?

# Discussion

# Acknowledgements

We are grateful to Claire Bergey, Ruthe Foushee, Ben Morris, and the members of the University of Chicago Chatter Lab and Northwestern University Child Language Lab for valuable discussion and feedback on this work.

# References

```{r}
# References will be generated automatically by Pandoc and included here.
# The following code is some latex to format the bibliography. Do not remove it.
```

\setlength{\parindent}{-0.1in} 
\setlength{\leftskip}{0.125in}
\noindent
