---
title: "Children's shift from CDS to ADS vocabulary across early childhood"
bibliography: CogSci2022-RegisterShift.bib
csl: apa6.csl
document-params: "10pt, letterpaper"

author-information: > 
    \author{{\large \bf Kennedy Casey (kbcasey@uchicago.edu)} 
    \AND {\large \bf Marisa Casillas (mcasillas@uchicago.edu)} \\Department of Comparative Human Development, University of Chicago \\ Chicago, IL 60637 USA}

abstract: >
    250-word abstract here
    
keywords: >
    child-directed speech; word production; linguistic input; social register; corpus analysis
    
output: cogsci2016::cogsci_paper
final-submission: \cogscifinalcopy
---

```{r global_options, include=FALSE}
knitr::opts_chunk$set(fig.width = 3, fig.height = 3, fig.crop = F, 
                      fig.pos = "tb", fig.path = 'figs/',
                      echo = F, warning = F, cache = F, 
                      message = F, sanitize = T)
```

```{r, libraries}
library(png)
library(grid)
library(xtable)
library(kableExtra)
library(tidyverse)
```

# Introduction

# Methods

## Corpus and item selection
We analyzed 8251 transcripts in the North American English collection of the Child Language Data Exchange System (CHILDES) database [@macwhinney2000childes]. The included transcripts were drawn from 52 individual corpora and featured 980 children up to 7 years of age (range = 1--84 months, _M_ = 33.5 months).

15 CDS/ADS word pairs were selected as test items based on their appearance on the MacArthur-Bates Communicative Development Inventory [@fenson1994variability] and their frequency of occurrence in CHILDES (criterion: at least 100 child-produced tokens and 100 other-produced tokens; see Table 1). 

- Cases where the same object, animal, or routine could be reasonably labeled with either form in typical communicative interactions with young children.

```{r freq_table, results = "asis"}
# source(get_freq.R)

freq <- read_csv("freq.csv")
cds_forms <- read_csv("../../data_prep/item_info.csv") %>%
    filter(form == "CDS") %>%
    pull(word)

ads_forms <- read_csv("../../data_prep/item_info.csv") %>%
    filter(form == "ADS") %>%
    pull(word)

cds_freq <- freq %>%
    filter(word %in% cds_forms)

ads_freq <- freq %>%
    filter(word %in% ads_forms)

freq <- bind_cols(cds_freq, ads_freq) %>%
    mutate(pair = paste0(word...1, '/', word...4)) %>%
    select(pair, child...2, other...3, child...5, other...6) 
colnames(freq) <- c("CDS/ADS word pair", "Child", "Other", "Child", "Other")

freq_table <- xtable::xtable(freq, digits = 0,
                             include.rownames = FALSE, 
                             caption = "CHILDES frequency for 15 CDS/ADS word pairs. Child-produced counts 
                             include tokens produced only by the target child. All other 
                             speakers' productions are included in the other-produced counts.") %>%
    xtable2kable(booktabs = TRUE, escape = FALSE) %>%
    kable_styling(font_size = 8) %>%
    add_header_above(c(" " = 2, "CDS tokens" = 2, "ADS tokens" = 2), bold = TRUE) %>%
    row_spec(0, bold = TRUE) %>%
    column_spec(2, italic = TRUE)

print(freq_table, include.rownames = FALSE,
      type = "latex", comment = FALSE, table.placement = "H")
```

# Results

# Discussion

# Acknowledgements
We are grateful to the members of the University of Chicago Chatter Lab and Northwestern University Child Language Lab for valuable discussion and feedback on this work.

# References

```{r}
```

\setlength{\parindent}{-0.1in} 
\setlength{\leftskip}{0.125in}
\noindent
